### 2020-05-17 abstract-old.tex

\begin{abstract}
\label{abstract}
In energy constraint wireless networks, radios usually operate in duty cycle mode. With low maintenance cost, asynchronous duty cycle management is widely adopted. Due to the diverse active schedules of different nodes, to achieve fast network flooding is challenge. Recently, concurrent flooding is a promising approach to speed up network flooding in asynchronous duty cycle networks. However, when all nodes are involved in flooding,
%Although it cannot miss any early delivery chances when all nodes are involved in flooding, the efficiency of concurrent flooding may be significantly degraded, especially in dense networks. The energy is also overused.
the key is an efficient approach to select a set of senders to simultaneously achieve fast concurrent flooding and efficient energy consumption. In this paper, we propose COFlood (Concurrent Opportunistic Flooding), which achieves the fast and energy-efficient concurrent flooding by opportunistically selecting senders based on an energy optimal flooding tree. The flooding tree structure provides full network coverage and keeps a low energy baseline. Then, with no extra control overhead, the opportunistic sender selection further filters out the additional early wake-up opportunities and bounds the number of total senders. We evaluate COFlood on two real testbeds. The results show both completion time and energy consumption are greatly reduced on the two testbeds.
\end{abstract}