### 2020-06-14 conclusion.tex
conclusion

\section{Conclusion}
\label{sec-conclusion}
In this work we designed a holistic system for shoulder surfing on smartphone, which serves as an up-to-date version of a threat model for shoulder surfing, and proved its efficiency. We proved that this threat towards screen privacy is imminent and can cause damage while undetected. It is our wish that this work can stur discussion in the field of screen privacy protection and proporgate defense mechanisms across critical mobile apps. Although this system proves to be highly efficient against unprotected screens, there are some simple methods to mitigate this unique threat while not cumbering the user. One of these defenses is displaying a dynamic background behind the characters, with a small protion of patterns colored similar to the texts. The constant changes will confuse most of the multi-frame SR algorithms, as they are often designed based on the assumption that all input images are reflections of the same scene. Other defence methods include active scanning for suspicious individuals, encrypted keyboards at password entry phases, etc.