### 2020-07-01 abstract.tex
修改“

\begin{abstract}
\label{abstract}
We use smartphones everywhere and privacy concerns have arisen for the ``shoulder surfing'' attack, where strangers peek at our phone in public areas. To mitigate this privacy threat, various countermeasures have been designed but mainly against attackers with the naked eye. With the development of smartphones and super resolution (SR) technique, a malicious attacker can peek from farther away with the assist of his/her smartphone camera and SR algorithms, rendering the underestimated threat model. To underline this concern, we present SRPeek, an end-to-end system of shoulder surfing deployed on commercial smartphones, including a unique deep neural network (DNN) architecture for multi-image SR. We implement SPPeek in Android. The experimental results demonstrate its efficiency and robustness, allowing a human to read 90\% of the characters at a distance of 1.8m with traditional lenses and 6m with optical zooming lenses. And it outperforms the state-of-the-arts and fills the blank space of multi-image SR for shoulder surfing attack.
\end{abstract}