### 2020-07-01 abstract.tex
intro

\begin{abstract}
\label{abstract}
We use smartphones everywhere and privacy concerns have arisen for the ``shoulder surfing'' attack, where strangers peek at our phone in public areas. To mitigate this privacy threat, various countermeasures have been designed but mainly against attackers with the naked eye. With the development of smartphones and super resolution (SR) techniques, a malicious attacker can peek from further away with the assist of the smartphone camera and SR algorithms, rendering the underestimated shoulder surfing threat model. To underline this concern, we present SRPeek, an end-to-end system of shoulder surfing deployed on commercial smartphones, including a unique deep neural network (DNN) for multi-image SR. We implement SPPeek in Android and conduct comprehensive experiments in controlled and real life scenarios. Evaluations demonstrate its efficiency and robustness, allowing a human to read 90\% of the characters at a distance of 1.8m with traditional lenses and 6m with optical zooming lenses. And it outperforms the state-of-the-arts and fills the blank space of multi-image SR for shoulder surfing attack.
\end{abstract}