### 2020-07-02 conclusion.tex
fix typos

\section{Conclusion}
\label{sec-conclusion}
In this work we designed a holistic system, SRPeek, for shoulder surfing on smartphone, which serves as an up-to-date version of a threat model for shoulder surfing, and proved its efficiency. We proved that this threat towards screen privacy is imminent and can steal critical information, including personal texts or passwords, from long distances, thus escaping detection. It is our wish that this work can stir discussion in the field of screen privacy protection and propagate defense mechanisms across critical mobile apps.

The core of SRPeek is a specially designed multi-frame SR network. With its innovative architecture this network outperforms other algorithms of the same field in our application. The design ideology enables this network to process higher levels of data integration ability while keeping a low calculation profile, and we believe the elements of this design can be used in other applications with large amounts of data, such as natural language processing or anomaly detection. Our model can also be used in OCR tasks when multiple images are available, functioning as a preprocessor to improve the quality of the images and increase accuracy.