### 2020-08-17 abstract.tex
newver

\begin{abstract}
\label{abstract}
We use smartphones everywhere and privacy concerns have arisen for the ``shoulder surfing'' attack, where strangers peek at our phone in public areas. To mitigate this privacy threat, various countermeasures have been designed but mainly against attackers with the naked eye. With the development of smartphones and super resolution (SR) techniques, a malicious attacker can peek from further away with the assist of his/her smartphone camera and SR algorithms, rendering the underestimated threat model. To underline this concern, we present \textsf{SRPeek}, an end-to-end system of shoulder surfing deployed on commercial smartphones, including a unique deep neural network (DNN) architecture for multi-image SR. We implement \textsf{SRPeek} in Android. The experimental results demonstrate its efficiency and robustness, allowing a human to read 90\% of the characters at a distance of 1.8m with traditional lenses and 6m with optical zooming lenses. And it outperforms the state-of-the-art and fills the blank space of multi-image SR for shoulder surfing attack.
\end{abstract}